Inference of Bifurcations with Differentiable Continuation

### 2020-07-27 docs/article.tex
Updates from Overleaf

\title{Inference and Synthesis of\\ co-dimension one Bifurcations}
\author{Gregory Szep$^{1,2}$, Neil Dalchau$^2$ and Attila Csikasz-Nagy$^1$}
\affil{$^1$King's College London, $^2$Microsoft Research Cambridge}
\date{\today} \maketitle

\section{Introduction and Motivation}
\begin{itemize}
\item qualitative equivalence more important
\item fitting to time courses not appropriate
\item gradient information makes optimisation tractable
\item similar works
\end{itemize}
\clearpage

\section{Problem Statement}
Suppose we parameterise a set of differential equations for states $u\in\Reals^N$ with a function $\rates$ in an unknown parameter space $\theta\in\Reals^M$. We would like these differential equations to obey a set of target bifurcations $\targets:=\{p_1\dots p_K\}$ along a known bifurcation parameter $p\in\Reals$. The following sections outline how a gradient descent algorithm could encourage predicted bifurcations $\predictions(\theta)$ coming from parameterised differential equations to match specified targets $\targets$. This would allow us to design differential equations using high-level qualitative constraints and sample qualitatively equivalent models in regions of optimal $\theta$. Let the differential equations be defined as
\begin{align}
	\partial_t u=\rates(z)
	\qquad\mathrm{where}\quad z:=(u,p)\quad
	\rates : \Reals^{N+1}\rightarrow\Reals^N
	\label{eq:model}
\end{align}
\section{Methodology}
First we view the steady states of this problem in terms of implicit space curves \cite{Goldman2005CurvatureSurfaces} in an augmented state space $z\in\Reals^{N+1}$ which combines the state $u$ and bifurcation parameter $p$. We then realise that the curvature of the Jacobian determinant $\Det(z):=|\partial\rates/\partial u|$ is maximised when bifurcations appear in the observed region of $z$. Luckily the curvature can be evaluated analytically. The steady state regions of $z$ are computed using a parameter continuation method \cite{Veltz2019PseudoArcLengthContinuation.jl,Farrell2016TheDiagrams}. Finally a piecewise-differentiable semi-supervised cost function $\loss(\theta)$ is proposed that would allow gradient descent algorithms to efficiently encourage bifurcations and match their locations to targets $\targets$.

For clarity we guide the reader through the methods with the following minimal models that explore the space of saddle-nodes $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ and pitchforks $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$. These minimal models are shown with targets $\targets$ in Figure \ref{fig:saddle-node} and Figure \ref{fig:pitchfork} respectively. These figures show that the curvature of the determinant $\Det(z)$ increases in the vicinity of bifurcations and crosses zero at the bifurcations. This gives us our first hint of what should be optimised to encourage bifurcations. Success on these minimal models demonstrates the proof of concept, and success in practice is demonstrated on more complicated genetic feedback systems taken from examples in synthetic biology.
\begin{figure}[H]
\centering{}
\captionsetup{justification=centering}
\includegraphics[width=11cm]{saddle-node}
\caption{Saddle-node model $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ with set $\theta=(2,-1)$ and targets $\targets=\{-1/2,1/2\}$. Lighter shades indicate the determinant crossing zero for unstable solutions}
\label{fig:saddle-node}
\end{figure}
\begin{figure}[H]
\centering{}
\captionsetup{justification=centering}
\includegraphics[width=11cm]{pitchfork}
\caption{Pitchfork model $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$ with set $\theta=(-1/10,-1)$ and target $\targets=\{0\}$. Lighter shades indicate the determinant crossing zero for unstable solutions}
\label{fig:pitchfork}
\end{figure}

@@

\begin{figure}[H]
\centering{}
\captionsetup{justification=centering}
\includegraphics[width=14cm]{tangent-field}
\caption{Left/Right : Calculated tangent fields $\tangent(z)$ for the saddle-node/pitchfork models for some set values of $\theta$ revealing that $\rates(z)=0$ picks out one of the traces in the field}
\label{fig:tangent-field}
\end{figure}

Figure \ref{fig:tangent-field} shows how the bifurcation curve defined by $\rates(z)=0$ picks out one of many traces in tangent field $\tangent(z)$ for the saddle and pitchfork. The tangent field $\tangent(z)$ can always be analytically evaluated by taking the determinant in \eqref{eq:tangent-field}. It becomes practical to proceed with calculations on $\tangent(z)$ in the whole space $z$ and pick out a single trace by solving $\rates(z)=0$ later. For the saddle-node and pitchfork respectively
\begin{align}
    \tangent(z)=\hat{u}-(\,3\theta_2 u^2+\theta_1\,)\,\hat{p}
    \qquad\qquad
    \tangent(z)=u\hat{u}-(\,3\theta_2 u^2+p\,)\,\hat{p}
    \label{eq:tangent-field-examples}
\end{align}

@@

Consequently the curvature of the determinant $\partial_{\tangent}^2\Det$ along the tangent field $\tangent(z)$ in the vicinity of bifurcations tends to increase. Fortunately this curvature can always be obtained in closed analytical form. The second directional derivative with respect to tangent field $\tangent(z)$
\begin{align}
    \partial_{\tangent}^2\Det &:=
    \bigg(
        \partial_z\left(
            \partial_z\Det \cdot \hat{\tangent}
        \right)
    \bigg)\cdot \hat{\tangent}\quad
    \mathrm{where} \quad \hat{\tangent}:=\frac{\tangent}{|\tangent|}\\
    &=
    \hat{\tangent}\cdot\partial_z^2\Det\cdot\hat{\tangent}
    \quad+\quad
    \partial_z\Det\cdot\partial_z\hat{\tangent}\cdot\hat{\tangent}
\end{align}
The first term in the expression above is the usual result involving the Hessian matrix $\partial_z^2\Det$ when taking second order directional derivatives. The second term involving the gradient $\partial_z\Det$ appears because $\hat{\tangent}(z)$ is a function of $z$ giving rise to a non-zero Jacobian matrix $\partial_z \hat{\tangent}$. Using expressions for tangent fields \eqref{eq:tangent-field-examples} and determinants for the saddle-node and pitchfork models the determinant curvatures respectively become
\begin{align}
    \partial_{\tangent}^2\Det =
    -\frac{6 \theta_2 \left(\theta _1^2-9 \theta _2^2 u^4+1\right)}
    {\left(\left(\theta _1+3 \theta _2 u^2\right){}^2+1\right){}^2}
    \qquad
    \partial_{\tangent}^2\Det =
    -\frac{2 u^2 \left(p+3 \theta _2 u^2\right) \left(9 \theta _2 \left(p+\theta _2 u^2\right)+1\right)}{
    ( \left(p+3 \theta _2 u^2\right)^2+u^2 )^2
    }
\end{align}
\subsection{Semi-supervised Cost Function}

@@

\begin{figure}[H]
\centering{}
\captionsetup{justification=centering}
\includegraphics[width=7cm]{likelihood-1.pdf}
\includegraphics[width=7cm]{likelihood-2.pdf}
\caption{Likelihood $L_{\theta}(z|\targets)$ shown in dark gold is sharply peaked around the bifurcation points $\predictions(\theta)$ and is maximal when they match the targets (Right). The light gold envelope shows the maxima of the likelihood as the points $\predictions(\theta)$ sweep across different values of $p$}
\label{fig:likelihood}
\end{figure}
Applying the gradient operator with respect to $\theta$ to the cost function, we find that not only the integrand but also the integration region needs to be differentiated. This requires the general Leibniz rule \cite{Flanders1973DifferentiationSign} which yields
\begin{align}
\frac{\partial\loss(\theta|\targets)}{\partial\theta}=
-\frac{
\int_{F=0}
  \mathrm{div}\!
\left(\frac{\partial z}{\partial\theta}
  \left(L+\lambda K^2\right)
\right)
+\frac{\partial L}{\partial\theta}+
2\lambda K\frac{\partial K}{\partial\theta}
\,\,\mathrm{d}z
}{
\int_{F=0}
\left(
  L+\lambda K^2
\right)\,\mathrm{d}z
}\label{eq:gradient}
\end{align}
where $\frac{\partial z}{\partial\theta}$ is the velocity field of the integration region due to changes in $\theta$. Fortunately the velocity field of implicitly defined regions can be solved for analytically \cite{Jos2011OnSurface}. This is done by taking the total derivative of the implicit equation
\begin{align}
    \frac{d\rates(z)}{d\theta}=\frac{\partial F}{\partial\theta}+
    \frac{\partial F}{\partial z}\frac{\partial z}{\partial\theta}
\end{align}
We can rearrange for $\frac{\partial z}{\partial\theta}$ using the Moore-Penrose inverse of the augmented Jacobian matrix $\frac{\partial F}{\partial z}$ which appeared in equation \eqref{eq:tangent-field}. Since the surface is defined by $\rates(z)=0$ we arrive at
\begin{align}
    \frac{\partial z}{\partial\theta} = - \frac{\partial F}{\partial z}^\top
    \left(\,
        \frac{\partial F}{\partial z}\,\frac{\partial F}{\partial z}^\top
    \right)^{-1}
    \frac{\partial F}{\partial\theta}
\end{align}
which for our two minimal models yield
\begin{align}
    \frac{\partial z}{\partial\theta} =
        \frac{ \hat{p}+\hat{u}\,(\,\theta_1+3\theta_2 u^2\,) }{1+(\,\theta_1+3\theta_2 u^2\,)^2}
        \left( u\,\hat{\theta}_1 + u^3\,\hat{\theta}_2 \right)
    \qquad
    \frac{\partial z}{\partial\theta} =
        \frac{ p\,\hat{u} + u\,(\,\hat{p}+3\theta_2u\,\hat{u}\,) }{(\,p+3\theta_2 u^2\,)^2 + u^2 }
        \left( \hat{\theta}_1 + u^3\,\hat{\theta}_2 \right)
    \label{eq:velocity-field-examples}
\end{align}
Thus the integrands in the gradient \eqref{eq:gradient} are completely analytic, leaving only region $\rates(z)=0$ to be evaluated numerically. Since these analytic forms can be pre-calculated, calculating the gradients shares the same computational complexity as calculating the steady states. 
\subsection{Basis Functions and Unsupervided Bias}

\subsection{Adaptive Hyperparameters}

\section{Method}

\subsection{Pseudo-arclength Continuation}
\label{sec:continuation}

\begin{itemize}
    \item predictor-corrector algorithm
    \end{itemize}

\subsubsection{Optimization details}

\begin{itemize}
    \item ADAM
\end{itemize}

\section{Normal Forms}
\label{sec:normal-forms}

\begin{itemize}
    \item saddle-node, pitchfork, transcritical
    \item benchmarks against other algos
\end{itemize}

\section{Chemical Reaction Networks}
\label{sec:networks}

\begin{itemize}
    \item toggle switch
    \item cell cycle (Attila)
    \item application to structure $\rightarrow$ function (Luca)
\end{itemize}

@@

\begin{itemize}
    \item hyperparam optimization
    \item hopf bifurcations
    \item pattern formation in pdes (Neil)
\end{itemize}

>>>>

\title{Inference of Bifurcations with Differentiable Continuation}
\author{Gregory Szep$^{1,2}$, Neil Dalchau$^2$ and Attila Csikasz-Nagy$^{1,3}$}
\affil{$^1$King's College London, UK, $^2$Microsoft Research Cambridge, UK, $^3$Pazmany Peter Catholic University, Budapest, Hungary}
\date{\today} \maketitle

\begin{abstract}
    In this work we propose a gradient-based semi-supervised approach for matching target bifurcations with parameterised differential equation models. The cost function contains a supervised likelihood term that is maximal when predicted bifurcations match the targets, and an unsupervised correction term that encourages bifurcations by maximising the curvature of the determinant of the Jacobian. The calculation of gradients with respect to parameters shares the same computational complexity as deflated pseudo-arclength continuation used to calculate the bifurcation diagram. We demonstrate model synthesis with minimal models which explore the space of saddle-node and pitchfork diagrams, and an applied model from synthetic biology. This work forms the basis for the inference of more complicated bifurcation diagrams which may include Andronov-Hopf points or extended to partial differential equations. 
\end{abstract}

\section{Introduction and Motivation}
\subsection{Background}
Backpropagation through differential equation solves has been a breakthrough over the past couple of years \cite{Chen2018NeuralEquations,Rackauckas2019DiffEqFlux.jl-AEquations} that enabled scalable parameter inference for differential equations. Optimisation targets, however, have mostly been expressed in the spatio-temporal domain. Acquisition of such data can be costly and can often contain over-constraining information generated by processing steps or the measurement device rather than the state variables that underpin the observed mechanism. In microscopy, for example, data is often reported in arbitrary fluoresce units allowing the observer to shift and scale data arbitrarily. Furthermore, such data may also not contain sufficient information about dynamical transients in order to identify kinetic parameters. The emerging picture suggests that identification of the qualitative dynamics -- the bifurcation diagram -- should precede any attempt at inferring kinetic parameters \cite{Stumpf2019ParameterBifurcations}.

In this work we propose a gradient-based semi-supervised approach that focuses on fitting qualitative dynamics, defined by state space structures, rather than kinetics. We find that calculating the gradients, which inform the optimisation of parameters, shares the same computational complexity as the algorithm used to calculate the bifurcation diagram. We use a predictor-corrector method called deflated pseudo-arclength continuation \cite{Farrell2016TheDiagrams,Veltz2019PseudoArcLengthContinuation.jl}. For partial differential equations, the computational complexity of calculating a single bifurcation diagram is not bounded since superpositions of localised solutions may give rise to uncountably many branches \cite{Avitabile2010ToEquation}. The complexity of computing a single branch, however, is bounded by the complexities of the chosen eigenvalue solver and corrector, and further decreased by adaptive stepping procedures \cite{Aruliah2016AlgorithmContinuation}.

We find that the cost function landscape contains basins that not only allow us to synthesise models with a desired bifurcation structure but also allow us to organise models in terms of topological and geometric equivalence. We discuss the relevance of this in model selection.

\subsection{Similar Works}

\begin{itemize}
    \item some juliacon submissions?
\end{itemize}
\clearpage

\section{Methodology}

\subsection{Problem Statement}
Suppose we parameterise a set of differential equations for states $u\in\Reals^N$ with a function $\rates$ in an unknown parameter space $\theta\in\Reals^M$. We would like these differential equations to obey a set of target bifurcations $\targets:=\{p_1\dots p_K\}$ along a known bifurcation parameter $p\in\Reals$. The following sections outline how a gradient descent algorithm could encourage predicted bifurcations $\predictions(\theta)$ coming from parameterised differential equations to match specified targets $\targets$. This would allow us to design differential equations using high-level qualitative constraints and sample qualitatively equivalent models in regions of optimal $\theta$. Let the differential equations be defined as
\begin{align}
	\partial_t u=\rates(z)
	\qquad\mathrm{where}\quad z:=(u,p)\quad
	\rates : \Reals^{N+1}\rightarrow\Reals^N
	\label{eq:model}
\end{align}

First we view the steady states of this problem in terms of implicit space curves \cite{Goldman2005CurvatureSurfaces} in an augmented state space $z\in\Reals^{N+1}$ which combines the state $u$ and bifurcation parameter $p$. We then realise that the curvature of the Jacobian determinant $\Det(z):=|\partial\rates/\partial u|$ is maximised when bifurcations appear in the observed region of $z$. Luckily the curvature can be evaluated analytically \cite{Goldman2005CurvatureSurfaces}. The steady state regions of $z$ are computed using a parameter continuation method \cite{Veltz2019PseudoArcLengthContinuation.jl,Farrell2016TheDiagrams}. Finally a differentiable semi-supervised cost function $\loss(\theta|\targets)$ is proposed that would allow gradient descent algorithms to efficiently encourage bifurcations and match their locations to targets $\targets$.

For clarity, we guide the reader through the methods with the following minimal models that explore the space of saddle-nodes $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ and pitchforks $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$. These minimal models are shown with targets $\targets$ in Figure \ref{fig:saddle-node} and Figure \ref{fig:pitchfork} respectively. These figures show that the curvature of the determinant $\Det(z)$ increases in the vicinity of bifurcations and crosses zero at the bifurcations. This gives us our first hint of what should be optimised to encourage bifurcations. Success on these minimal models demonstrates the proof of concept, and success in practice is demonstrated on a four component genetic toggle switch model taken from synthetic biology.
\begin{figure}[H]
\centering{}
\captionsetup{justification=centering}
\includegraphics[width=11cm]{saddle-node}
\caption{Saddle-node model $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ with set $\theta=(2,-1)$ and targets $\targets=\{-1/2,1/2\}$. Lighter shades indicate the determinant crossing zero for unstable solutions}
\label{fig:saddle-node}
\end{figure}
\begin{figure}[H]
\centering{}
\captionsetup{justification=centering}
\includegraphics[width=11cm]{pitchfork}
\caption{Pitchfork model $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$ with set $\theta=(-1/10,-1)$ and target $\targets=\{0\}$. Lighter shades indicate the determinant crossing zero for unstable solutions}
\label{fig:pitchfork}
\end{figure}

@@

Figure \ref{fig:determinant-curvature} shows how the bifurcation curve defined by $\rates(z)=0$ picks out one of many traces in tangent field $\tangent(z)$ for the saddle and pitchfork. The tangent field $\tangent(z)$ can always be analytically evaluated by taking the determinant in \eqref{eq:tangent-field}. We will proceed with calculations on $\tangent(z)$ in the whole space $z$ and pick out a single trace by solving $\rates(z)=0$ later. For our two models
\begin{align}
    \underset{\mathrm{saddle-node\,\,model}}{
    \tangent(z)=\hat{u}-(\,3\theta_2 u^2+\theta_1\,)\,\hat{p}}
    \qquad\qquad
    \underset{\mathrm{pitchfork\,\,model}}{
    \tangent(z)=u\hat{u}-(\,3\theta_2 u^2+p\,)\,\hat{p}}
    \label{eq:tangent-field-examples}
\end{align}

@@

Consequently the curvature of the determinant $\partial_{\tangent}^2\Det$ along the tangent field $\tangent(z)$ in the vicinity of bifurcations tends to increase. Fortunately this curvature can always be obtained in closed analytical form. The second directional derivative with respect to tangent field $\tangent(z)$
\begin{align}
    \partial_{\tangent}^2\Det &:=
    \bigg(
        \partial_z\left(
            \partial_z\Det \cdot \hat{\tangent}
        \right)
    \bigg)\cdot \hat{\tangent}\quad
    \mathrm{where} \quad \hat{\tangent}:=\frac{\tangent}{|\tangent|}\\
    &=
    \hat{\tangent}\cdot\partial_z^2\Det\cdot\hat{\tangent}
    \quad+\quad
    \partial_z\Det\cdot\partial_z\hat{\tangent}\cdot\hat{\tangent}
\end{align}
The first term in the expression above is the usual result involving the Hessian matrix $\partial_z^2\Det$ when taking second order directional derivatives. The second term involving the gradient $\partial_z\Det$ appears because $\hat{\tangent}(z)$ is a function of $z$ giving rise to a non-zero Jacobian matrix $\partial_z \hat{\tangent}$. Using expressions for tangent fields \eqref{eq:tangent-field-examples} and determinants for the saddle-node and pitchfork models the determinant curvatures respectively become
\begin{align}
    \underset{\mathrm{saddle-node\,\,model}}{
    \partial_{\tangent}^2\Det =
    -\frac{6 \theta_2 \left(\theta _1^2-9 \theta _2^2 u^4+1\right)}
    {\left(\left(\theta _1+3 \theta _2 u^2\right){}^2+1\right){}^2}}
    \qquad
    \underset{\mathrm{pitchfork\,\,model}}{
    \partial_{\tangent}^2\Det =
    -\frac{2 u^2 \left(p+3 \theta _2 u^2\right) \left(9 \theta _2 \left(p+\theta _2 u^2\right)+1\right)}{
    ( \left(p+3 \theta _2 u^2\right)^2+u^2 )^2
    }}
\end{align}
\subsection{Semi-supervised Cost Function}

@@

\begin{figure}[H]
\centering{}
\captionsetup{justification=centering}
\includegraphics[width=7cm]{likelihood-1.pdf}
\includegraphics[width=7cm]{likelihood-2.pdf}
\caption{Likelihood $L_{\theta}(z|\targets)$ shown in dark gold is sharply peaked around the bifurcation points $\predictions(\theta)$ and is maximal when they match the targets (Right). The light gold envelope shows the maxima of the likelihood as the points $\predictions(\theta)$ sweep across different values of $p$}
\label{fig:likelihood}
\end{figure}
Applying the gradient operator with respect to $\theta$ to the cost function, we find that not only the integrand but also the integration region needs to be differentiated. This requires the general Leibniz rule \cite{Flanders1973DifferentiationSign} which yields
\begin{align}
\frac{\partial\loss(\theta|\targets)}{\partial\theta}=
-\frac{
\int_{F=0}
  \mathrm{div}\!
\left(\frac{\partial z}{\partial\theta}
  \left(L+\lambda K^2\right)
\right)
+\frac{\partial L}{\partial\theta}+
2\lambda K\frac{\partial K}{\partial\theta}
\,\,\mathrm{d}z
}{
\int_{F=0}
\left(
  L+\lambda K^2
\right)\,\mathrm{d}z
}\label{eq:gradient}
\end{align}
where $\frac{\partial z}{\partial\theta}$ is the velocity field of the integration region due to changes in $\theta$. Fortunately the velocity field of implicitly defined regions can be solved for analytically \cite{Jos2011OnSurface}. This is done by taking the total derivative of the implicit equation
\begin{align}
    \frac{d\rates(z)}{d\theta}=\frac{\partial F}{\partial\theta}+
    \frac{\partial F}{\partial z}\frac{\partial z}{\partial\theta}
\end{align}
We can rearrange for $\frac{\partial z}{\partial\theta}$ using the Moore-Penrose inverse of the augmented Jacobian matrix $\frac{\partial F}{\partial z}$ which appeared in equation \eqref{eq:tangent-field}. Since the surface is defined by $\rates(z)=0$ we arrive at
\begin{align}
    \frac{\partial z}{\partial\theta} = - \frac{\partial F}{\partial z}^\top
    \left(\,
        \frac{\partial F}{\partial z}\,\frac{\partial F}{\partial z}^\top
    \right)^{-1}
    \frac{\partial F}{\partial\theta}
\end{align}
which for our two minimal models yield
\begin{align}
    \underset{\mathrm{saddle-node\,\,model}}{
    \frac{\partial z}{\partial\theta} =
        \frac{ \hat{p}+\hat{u}\,(\,\theta_1+3\theta_2 u^2\,) }{1+(\,\theta_1+3\theta_2 u^2\,)^2}
        \left( u\,\hat{\theta}_1 + u^3\,\hat{\theta}_2 \right)}
    \qquad
    \underset{\mathrm{pitchfork\,\,model}}{
    \frac{\partial z}{\partial\theta} =
        \frac{ p\,\hat{u} + u\,(\,\hat{p}+3\theta_2u\,\hat{u}\,) }{(\,p+3\theta_2 u^2\,)^2 + u^2 }
        \left( \hat{\theta}_1 + u^3\,\hat{\theta}_2 \right)}
    \label{eq:velocity-field-examples}
\end{align}
Thus the integrands in the gradient \eqref{eq:gradient} are completely analytic, leaving only region $\rates(z)=0$ to be evaluated numerically. Since these analytic forms can be pre-calculated, calculating the gradients shares the same computational complexity as calculating the steady states. 
% \subsection{Basis Functions and Unsupervided Bias}
% \subsection{Pseudo-arclength Continuation}

\subsection{Optimisation Loop}
We use \texttt{BifurcationKit.jl} \cite{Veltz2019PseudoArcLengthContinuation.jl} to calculate bifurcation diagrams, which requires the rate function $\rates$, its Jacobian $J_\theta$, and an initial guess $u_0$ at an initial parameter $p_0$. Let this algorithm be called by function $\mathrm{getSteadyStates}$ and return the steady states that form the integration region $\Omega$ and a new guess $u_0$ that is the true steady state at $p_0$.

The algorithm also takes hyperparameters $\beta$ which will contain arclength step sizes, eigenvalue and Newton solver options and unsupervised correction factor $\lambda$. The adaptive hyperparameter update is done by $\mathrm{updateHyperparameters}$ and should depend on the current solutions $\Omega$ and the targets $\targets$. The step sizes and solver options should adapt to the expected space of the solutions $\Omega$ so that the waiting time between iterations is minimised. The unsupervised correction factor $\lambda$ should be large when there are no predicted bifurcations and decrease as the number of predicted bifurcations approaches the number of targets.

The evaluation of the cost gradient \eqref{eq:gradient}

\\
\begin{algorithm*}[H]
\label{alg:optimisation-loop}
\SetAlgoLined
\textbf{Inputs} Function $F_{\theta}:\mathbb{R}^{N+1}\rightarrow\mathbb{R}^{N}$ for $u\in\mathbb{R}^N$ and $p\in\mathbb{R}$\\ target bifurcations $\mathcal{D}=\{p:p\in[p_{\min},p_{\max}]\}$ and convergence tolerance $\varepsilon$\\
\textbf{Output} Optimised $\theta\in\mathbb{R}^M$ that satisfy targets $\targets$\\
$J_{\theta}\leftarrow\partial F/\partial u$\qquad\,
$\Delta_{\theta}\leftarrow|\mathrm{det}\,J_{\theta}|$\qquad$K_{\theta}\leftarrow\partial^2\Delta_{\theta}/\partial\tangent^2$\\
$\frac{\partial z}{\partial\theta}\leftarrow- \frac{\partial F}{\partial z}^\top
    \left(\,
        \frac{\partial F}{\partial z}\,\frac{\partial F}{\partial z}^\top
    \right)^{-1}
    \frac{\partial F}{\partial\theta}$\\
$u_0\leftarrow\mathrm{rand}\in\mathbb{R}^{N}$\quad\,$\theta\leftarrow\mathrm{rand}\in\mathbb{R}^{M}$\\
$\beta\leftarrow\mathrm{getHyperparameters}(\targets)$\\
\While{$\mathrm{tolerance}(\beta)>\varepsilon$}{
  $\Omega,u_0\leftarrow\mathrm{getSteadyStates}(F_{\theta},J_{\theta},u_0,\beta)$\\
  $\beta\leftarrow\mathrm{updateHyperparameters}(\beta,\Omega,\targets)$\\
  $\partial L\leftarrow\mathrm{costGradient}(\Delta_{\theta},K_{\theta},V_{\theta},\Omega,\targets)$\\
  $\theta\leftarrow\mathrm{updateParameters}(\theta,\partial L,\beta)$
}
\caption{Bifurcation Optimisation Loop}
\end{algorithm*}

\section{Results}

\subsection{Model Synthesis}

\subsection{Model Selection}

@@

\begin{itemize}
    \item cell cycle / hopf
    \item application to structure $\rightarrow$ function (Luca)
    \item pattern formation in pdes
\end{itemize}
